### 2021-02-15 main/intro.tex
rev03

\par Early algorithms for DOA estimation were based on the beamforming concept \cite{krim1996two}, where the array's reception beam is steered (mechanically or electronically) to multiple directions in search of received energy peaks which are interpreted as valid emitters.
One notable improvement to the convectional beamformer \cite{van2004optimum} was Capon's beamformer  (aka MVDR in acoustics related literature) which attempts to minimize the power contributed by noise and any signals coming from other directions than the DOA of interest, while maintaining a fixed gain
in the "look direction".

@@

MUSIC (Multiple SIgnal Classification) algorithm \cite{schmidt1986multiple} which decomposes the estimated covariance matrix to its signal and noise eigenvectors, using orthogonal projection assuming that the noise space is orthogonal to the array manifold.
It is interesting to note that while earlier works were mostly derived in the context of time series analysis and later applied to the sensor array problem, MUSIC was indeed originally presented as a DOA estimator.
\par A Less practical, but optimal in the sense of RMSE \cite{krim1996two}, is the statistical based ML approach which tries to find the most probable DOA assuming a known PDF.

@@

Adding the fact that sub-optimal methods such as MUSIC were proven achieve the CRB under some practical assumptions (e.g. non-coherent signals) \cite{stoica1989music}, such methods (i.e. the sub optimal estimators) are substantially more common in practice. 

@@

One approach, commonly referenced as ``virtual arrays'' \cite{pal2010nested,chevalier2005virtual,dogan1995applications} deals with the extraction of samples originated in sensors that do not really exist by using high (higher than 2) order statistics and manipulating multiple statistical cross-terms in order to estimate statistical characteristics of signals impinging in missing sensors.

@@

As in many adaptive processing applications, the two methods rely on steepest descend optimization, which expose the adaptive processing approach to errors related to optimization parameters choices such step size etc.

@@

Inspired by this analogy, which is thoroughly discussed in Sec.~\ref{sec:prlm_FIR_IIR}, we were motivated to find the spatial counterpart of the temporal IIR filter, both from academic curiousness and the known advantages of the IIR filter's frequency (DOA by analogy) response, hoping to gain spatial performance by finding the right architecture.

@@

     \caption{Referring to consecutive sub arrays' output as sequential output of the same ULA.}

@@

Obviously, the former approach is very sensitive to the synthetic delay accuracy and involves temporal domain processing which we try to avoid in this work to achieve purely spatial processing.  

@@

\par Other interesting works \cite{Hum2009BeamformingFilters,madanayake2008speed} in the field of ultra-wideband (UWB) treat the spatial and temporal diversity, which reside in multiple consecutive snapshots of the array output, as if the signal is sampled in two independent temporal frequencies.
Inspired by \cite{bruton1985three}, the authors show that impinging plane waves are represented as straight lines tilted according to their DOA and suggest methods to design line filters in the spatio-temporal plane.
In \cite{Hum2009BeamformingFilters}, the authors also address the fact that ideal straight line filter may not be achieved due to the fact that the filter is periodic in both frequencies, causing the 2D frequency response the ``bend`` close to the sampling frequencies (see Fig.~\ref{fig_intro_UWB_bend}).

@@

\par The novelty, compared to traditional array processing, is the incorporation of spatial feedback, which we prove to be the spatial domain equivalent of temporal domain IIR filtering.

@@

\par In the followings, we elaborate in Chapter~\ref{chap:prelims} on some relevant theoretic concepts which serve as the foundation of this presentation.
Then our novel architecture is presented and analysed in Chapter.~\ref{chap:firstchap} followed by a thorough concluding discussion on the findings and future research in Chapter.~\ref{chap:futureResearch}.

>>>>

\par Early algorithms for direction of arrival (DOA) estimation were based on the beamforming concept \cite{krim1996two}, where the array's reception beam is steered (mechanically or electronically) to multiple directions in search of received energy peaks which are interpreted as valid emitters.
One notable improvement to the convectional beamformer \cite{van2004optimum} was Capon's beamformer  (aka minimum variance distortionless response (MVDR) in acoustics related literature) which attempts to minimize the power contributed by noise and any signals coming from other directions than the DOA of interest, while maintaining a fixed gain
in the desired direction.

@@

MUSIC (Multiple SIgnal Classification) algorithm \cite{schmidt1986multiple} which decomposes the estimated covariance matrix to its signal and noise eigenvectors, using orthogonal projection exploiting the fact that the noise space is orthogonal to the signal subspace.
It is interesting to note that while earlier works were mostly derived in the context of time series analysis and only later applied to the sensor array problems, MUSIC was originally presented as a DOA estimator.
\par A less practical, but optimal in the sense of RMSE \cite{krim1996two}, is the maximum likelihood (ML) approach which tries to find the most probable DOA assuming a known PDF.

@@

Adding the fact that sub-optimal methods such as MUSIC were proven to achieve the CRB under some practical assumptions (e.g. non-coherent signals) \cite{stoica1989music}, these sub optimal estimators are substantially more common in practice. 

@@

One approach, commonly referenced as ``virtual arrays'' \cite{pal2010nested,chevalier2005virtual,dogan1995applications} deals with the extraction of samples originated in sensors that do not really exist, by using high (higher than 2) order statistics and manipulating multiple statistical cross-terms in order to estimate statistical characteristics of signals impinging in missing sensors.

@@

As in many adaptive processing applications, the two methods rely on steepest descend optimization, which exposes, the adaptive processing approach to errors related to optimization parameters choices such step size etc.

@@

Inspired by this analogy, which is thoroughly discussed in Sec.~\ref{sec:prlm_FIR_IIR}, we were motivated to find the spatial counterpart of the temporal IIR filter, both from academic curiousness and due to the known advantages of the IIR filter's performance.

@@

     \caption{In \cite{wen2013extending} subsections on a single large ULA are used to approximate an IIR response}

@@

Obviously, the former approach is very sensitive to the synthetic delay accuracy and involves temporal domain processing (which we try to avoid in this work in order to achieve purely spatial processing).  

@@

\par Other interesting works \cite{Hum2009BeamformingFilters,madanayake2008speed} in the field of ultra-wideband (UWB) signals treat the spatial and temporal diversity, which reside in multiple consecutive snapshots of the array output, as if the signal is sampled in two independent temporal frequencies.
Inspired by \cite{bruton1985three}, the authors show that impinging plane waves are represented as straight lines tilted according to their DOA and suggest methods to design line filters in that spatio-temporal plane.
In \cite{Hum2009BeamformingFilters}, the authors also address the fact that ideal straight line filter is mathematically unachievable due to spatio-temporal frequencies periodicity.
This issue manifests as a ``bend`` close to the sampling frequencies (see Fig.~\ref{fig_intro_UWB_bend}).

@@

\par The novelty, compared to traditional array processing, is by incorporating spatial feedback, which we prove to be the spatial domain equivalent of temporal domain IIR filtering.

@@

\par In the followings, we elaborate in Chapter~\ref{chap:prelims} on some relevant theoretic concepts which serve as the foundation of this work.
Then, our novel architecture is presented and analysed in Chapter.~\ref{chap:firstchap} followed by a thorough concluding discussion of the findings and future research directions in Chapter.~\ref{chap:conclusion}.

### 2021-02-15 main/prelims.tex
rev03

\chapter{Preliminaries}
\label{chap:prelims}
As a preliminary discussion to the presentation of our contribution, we overview some the key theoretical basis of the thesis.
First, a short overview of array-processing is presented in Sec.~\ref{sec:prlm_sensorArrays}, accompanied by a complementary wave propagation discussion in Sec.~\ref{sec:prlm_propWaveField}.
To accommodate the discussion for spatial filtering, we also mention the time-space relation of spatially sampled signals in Sec.~\ref{sec:prlm_timeSpaceSig}.
In Sec.~\ref{sec:prlm_localization}, we further focus the spatial filtering discussion to localization and overview some of its fundamentals.
Following the FIR/IIR filters related discussion in Chapter.~\ref{chap:intro}, we further elaborate and compare the two architectures in Sec.~\ref{sec:prlm_FIR_IIR}.
To complete the basics overview, we also discuss array performance measuring in Sec.~\ref{sec:prlm_array_perf} following the classical localization related literature.
% \begin{definition}
% The \emph{von Neumann model} of a computer, also known as the \emph{Princeton architecture} is an architecture for digital computers, which consists of a processing units, containing an ALU and processing registers; a control unit consisting of an instruction register and a program counter; a memory unit which stores both data and instructions; and input-and-output mechanisms.
% \end{definition}

>>>>

\chapter{Preliminaries}
\label{chap:prelims}
As a preliminary discussion to the presentation of our contribution, we overview some the key theoretical basis of the thesis.
First, a short overview of array-processing is presented in Sec.~\ref{sec:prlm_sensorArrays}, accompanied by a complementary wave propagation discussion in Sec.~\ref{sec:prlm_propWaveField}.
To accommodate the discussion for spatial filtering, we also mention the time-space relation of spatially sampled signals in Sec.~\ref{sec:prlm_timeSpaceSig}.
In Sec.~\ref{sec:prlm_localization}, we further focus the spatial filtering discussion to localization and overview some of its fundamentals.
Following the FIR/IIR filters related issue, which was previously mentioned in Chapter.~\ref{chap:intro}, we further elaborate and compare the two architectures in Sec.~\ref{sec:prlm_FIR_IIR}.
To complete the basics overview, we also discuss array performance measures in Sec.~\ref{sec:prlm_array_perf} following classical array processing literature.
% \begin{definition}
% The \emph{von Neumann model} of a computer, also known as the \emph{Princeton architecture} is an architecture for digital computers, which consists of a processing units, containing an ALU and processing registers; a control unit consisting of an instruction register and a program counter; a memory unit which stores both data and instructions; and input-and-output mechanisms.
% \end{definition}
